### 2018-07-16 misc/thesis-hardbound/Appendices/AppendixA.tex
Update 16/7/2018

% Appendix A

\chapter{Calculating Sun Position from Earth's Surface} % Main appendix title
\label{append:sunposition} % For referencing this appendix elsewhere, use \ref{append:motor}

\lhead{Appendix A. \emph{Calculating Sun Position}} % This is for the header on each page - perhaps a shortened title

The position of the sun from the earth's surface is generally measured using two entities---elevation angle and zenith angle. The angle projected from the sun with reference to the horizon is called the elevation angle. The compass angle of the sun from the $0^{\circ}$ N direction is called the azimuth angle. We use the Jean Meeus algorithm~\cite{SPA} that has an error of $\pm 3\cdot 10^{-4}$ degrees. This is one of the most popular algorithms, as its implementation is available in various platforms. It consists of an online web version~\footnote{\url{https://www.nrel.gov/midc/solpos/spa.html}}; also it is available in Python~\footnote{\url{http://pysolar.org/}} and MATLAB~\footnote{\url{http://pvpmc.org/pv-lib/}} programming languages.