### 2018-07-16 misc/thesis-hardbound/Misc/abbreviations.tex

\chapter*{\centering Abbreviations}
\begin{longtable}[l]{p{100pt}p{400pt}}
\textbf{AEB} & Automatic Exposure Bracketing  \\
\textbf{BOW} & Bag-Of-Words \\
\textbf{CDF} & Cumulative Distribution Plot \\
\textbf{CLAHE} & Contrast-Limited Adaptive Histogram Equalization \\
\textbf{colorHIST} & Color Histogram \\
\textbf{CRE} & Cloud Radiative Effect \\
\textbf{DBSCAN} & Density-Based Spatial Clustering of Applications with Noise \\
\textbf{DR} & Dynamic Range\\
\textbf{dSIFT} & Dense Scale Invariant Feature Transform \\
\textbf{EV} & Exposure Value \\
\textbf{FF} & Fractal Feature \\
\textbf{FFT} & Fast Fourier Transform \\
\textbf{FM} & Fourier Magnitude \\
\textbf{FN} & False Negative \\
\textbf{FP} & False Positive \\
\textbf{GEOSS} & Global
Earth Observation System of Systems \\
\textbf{GHI} & Global Horizontal Irradiance \\
\textbf{GLCM} & Gray-Level Co-occurrence Matrix \\
\textbf{GLDS} & Gray Level Difference Statistics \\
\textbf{GPS} & Global Positioning System \\
\textbf{HDR} & High Dynamic Range \\
\textbf{HDRI} & High Dynamic Range Imaging\\
\textbf{HF} & Heinle Feature \\
\textbf{HYTA} & Hybrid Thresholding Algorithm \\
\textbf{ICESat} & Ice, Cloud, and land Elevation Satellite \\
\textbf{ICLTP} & Illumination-invariant Completed Local Ternary Pattern \\
\textbf{IR} & Infra-Red \\
\textbf{KL} & Kullback-Leibler \\
\textbf{LBP} & Local Binary Patterns \\
\textbf{LCAV} & Audiovisual Communications Laboratory \\
\textbf{LDR} & Low Dynamic Range \\
\textbf{LED} & Light-Emitting Diode \\
\textbf{LiDAR} & Light Detection And Ranging \\
\textbf{MODIS} & Moderate-Resolution Imaging Spectroradiometer \\
\textbf{MSS} & Meteorological Service Singapore \\
\textbf{NASA} & National Aeronautics and Space Administration \\
\textbf{NIR} & Near-Infra-Red \\
\textbf{OC} & Operating Characteristics \\
\textbf{PBI} & Pearson's Bimodality Index \\
\textbf{PCA} & Principal Component Analysis \\
\textbf{PLS} & Partial Least Squares \\
\textbf{PV} & Photo Voltaic \\
\textbf{RMSE} & Root Mean Square Error \\
\textbf{ROC} & Receiver Operating Characteristic \\
\textbf{SDK} & Software Development Kit \\
\textbf{SGLDM} & Spatial Gray Level Dependence Matrices \\
\textbf{SHARP} & Sequential Half space Aggregation for Reconstructing a Point \\
\textbf{SIFT} & Scale Invariant Feature Transform \\
\textbf{SLIC} & Simple Linear Iterative Clustering \\
\textbf{SVM} & Support Vector Machine \\
\textbf{SWIMCAT} & Singapore Whole sky IMaging CATegories database \\
\textbf{SWIMSEG} & Singapore Whole sky IMaging SEGmentation database \\
\textbf{TN} & True Negative \\
\textbf{TP} & True Positive \\
\textbf{USB} & Universal Serial Bus \\
\textbf{WAHRSIS} & Wide Angle High Resolution Sky Imaging System \\
\textbf{WMO} & World Meteorological Organization \\
\textbf{WSI} & Whole Sky Imager \\
\textbf{YES} & Yankee Environmental Systems\\
\end{longtable}