### 2018-07-16 misc/thesis-hardbound/Misc/ack.tex

\chapter*{\centering Acknowledgment}

I would like to offer sincere gratitude to my two wonderful supervisors---Prof.\ Lee Yee Hui \& Dr.\ Stefan Winkler. Thanks Prof.\ Lee for admitting me into your research group, and for introducing me into the wonderful world of research and academics. You are a passionate professor, a dedicated researcher and a wonderful friend---sincere thanks for believing in me always, especially during the formative years of my PhD and course work. A huge thanks to Stefan for all your support and guidance throughout my PhD---this thesis is in its current form because of the numerous discussions and regular meetings with you. Thank you for introducing me into the world of \emph{computer vision}, for meticulous reading and editing the papers, and for guiding me throughout this journey.

I have been extremely fortunate to be a visiting Doctoral Assistant at Audiovisual Communication Laboratory (LCAV), at \'{E}cole Polytechnique F\'{e}d\'{e}rale de Lausanne, Switzerland, under the supervision of Prof.\ Martin Vetterli and Dr.\ Adam Scholefield, during my PhD. Thank you Martin for this wonderful opportunity---it was my honor to be a part of LCAV and interact with the brilliant minds. I sincerely appreciate your guidance throughout the period, and always inspiring me to be a better researcher. A special shout-out to Adam for your time and continuous support, for all the white-board discussions at \emph{BC 366}, and for providing me all the confidence.

Friends have been an integral part of this PhD journey, and I am truly indebted to them for their love and support. I would like to thank---Prof.\ J.\ T.\ Ong, Dr.\ Meng Yusong, Dr.\ Lakshmi Sutha, Ma Ting, Mao Xiao Hong, Yeo Jun Xiang, Ian Bajaj, Mi Siya, Shilpa Manandhar (thanks for the warm friendship and companionship) and Yuan Feng (thanks for the wonderful times). A warm thanks to Ong Kee Wang, Darko Jokic \& Charles Wong for assisting to build the mechanical models of our sky cameras. And of course, Florian Savoy---a true engineer and a wonderful friend; thanks for all the co-authorships, \emph{nuits blanches} before the conference deadlines and for all our technical discussions. \emph{Merci Florian!}

I would also like to thank all my friends I met during my internship at Lausanne. Thanks to Alireza Ghasemi, Mitra Fatemi, Runwei Zhang for being a wonderful office-mate, Niranjan Thanikachalam for all the guidance and friendship, Gilles Baechler for the friendship (and running sessions by the lake), Marta Martinez, and of course, Julien Reichel for being an amazing host. Special thanks to Robin Scheibler for introducing me to \emph{python}, for the numerous discussions over coffee, and for being a wonderful friend. A huge thanks to Radmila Popovic for all the love and support. And finally, special thanks to my dearest friend Miranda Krekovi{\'c} for all the good times spent at Lausanne and beyond---I sincerely appreciate and respect it. \emph{Puno hvala Miranda!}

Thank you Subhrajit, Sumanta, Zishan, Binit, Sana, Ankush, Deval, Sushant and my wonderful flatmates---Jithin, Dhirendra, Pravesh \& Raunaq (who were no less than my brothers) for all the love and support throughout this journey. I'm sure I have missed mentioning the names of many who were beside me at all the happy and sad times. Singapore is beautiful because of your presence!

I would also like to thank my thesis advisory committee members Prof.\ Wang Gang and Prof.\ Cai Jianfei for advising this thesis work. Also, thanks to my PhD oral examination panel members, Prof.\ Bi Guoan, Prof.\ Mohammed Yakoob Siyal and Prof.\ Anamitra Makur for taking their respective time in reading this thesis, and providing valuable feedback on the same. Sincere thanks to the anonymous internal and external PhD thesis examiners, who were amongst the first to closely examine the thesis and provide useful feedback.

I would also like to acknowledge the financial support from Nanyang Technological University Singapore and Ministry of Defence (MINDEF) Singapore; and administrative support from Graduate Programme Office, School of Electrical and Electronic Engineering. Thanks Christina Wong for your continuous help and support, and your infectious laughter definitely made the experience more enjoyable. Special thanks to Tay TingFang for assisting me throughout the process of my thesis examination (and for acceding to all my technical and logistical requests for my thesis examination).

Thanks to Dr.\ John Hammersley, Shelly Miller \& Dr.\ Lim Lian Tze from Overleaf for sharing the vision of open science with the community. This entire thesis was written in \emph{Overleaf}; and it was my pleasure to spread the \emph{Overleaf} joy amongst the NTU community, as an advisor.

Finally, I would like to offer my sincere thanks and gratitude to my parents Ashoka Dev \& Bidhu Bhusan Dev, my brother Bedabrata Dev and my sister-in-law Rupsa Banerjee. Thank you for your unconditional love and support and for the innumerable sacrifices. Special thanks to my mother who inspired me to pursue a doctoral degree at the very outset!