### 2018-07-16 misc/thesis-hardbound/Misc/pubs.tex

\chapter*{\centering Author's Publications}

\lhead{Author's Publications}

\textbf{Journals}
\begin{itemize}[leftmargin=*]
\item S. Dev, F. M. Savoy, Y. H. Lee, S. Winkler, Rough-set Based Visible Color Channel Selection, \emph{IEEE Geoscience and Remote Sensing Letters}, vol. 14, no. 1, pp. 52--56, Jan 2017. 
\item S. Dev, Y. H. Lee, S. Winkler, Color-based segmentation of sky/cloud images from ground-based cameras, \emph{IEEE Journal of Selected Topics in Applied Earth Observations and Remote Sensing}, vol. 10, no. 1, pp. 231--242, Jan 2017. 
\item S. Dev, B. Wen, Y. H. Lee, S. Winkler, Machine Learning Techniques and Applications For Ground-based Image Analysis, \emph{IEEE Geoscience and Remote Sensing Magazine, Special Issue on Advances in Machine Learning for Remote Sensing and Geosciences}, vol. 4, no. 2, pp. 79--93, Jun 2016. 
\end{itemize}

\noindent 
\textbf{Conferences}
\begin{itemize}[leftmargin=*]
\item S. Dev, F. M. Savoy, Y. H. Lee and S. Winkler, Nighttime sky/cloud image segmentation, \emph{IEEE International Conference on Image Processing (ICIP)}, 2017.
\item F. Savoy, S. Dev, Y. H. Lee and S. Winkler, Stereoscopic Cloud Base Reconstruction Using High-Resolution Whole Sky Imagers, \emph{IEEE International Conference on Image Processing (ICIP)}, 2017. 
\item S. Dev, S. Manandhar, F. Yuan, Y. H. Lee and S. Winkler, Cloud Radiative Effect Study Using Sky Camera, \emph{Proc. IEEE AP-S Symposium on Antennas and Propagation and USNC-URSI Radio Science Meeting}, 2017.
\item  S. Dev, S. Manandhar, Y. H. Lee and S. Winkler,  Detecting rainfall onset using sky images, \emph{TENCON 2016 - 2016 IEEE Region 10 Conference}, 2016.
\item  S. Dev, F. M. Savoy, Y. H. Lee and S. Winkler, Short-term prediction of localized cloud motion using ground-based sky imagers, \emph{TENCON 2016 - 2016 IEEE Region 10 Conference}, 2016.
\item  S. Manandhar, F. Yuan, S. Dev, Y. H. Lee and Y. S. Meng, Weather radar to detect cloud occurence level, \emph{TENCON 2016 - 2016 IEEE Region 10 Conference}, 2016.
\item  S. Dev, F. M. Savoy, Y. H. Lee, S. Winkler, Estimation of solar irradiance using ground-based whole sky imagers, \emph{IEEE International Geoscience and Remote Sensing Symposium (IGARSS)}, 2016.
\item F. M. Savoy, S. Dev, Y. H. Lee, S. Winkler, Geo-referencing and stereo calibration of ground-based whole sky imagers using the sun trajectory, \emph{IEEE International Geoscience and Remote Sensing Symposium (IGARSS)}, 2016.
\item S. Manandhar, Y. H. Lee, S. Dev, GPS Derived PWV for rainfall monitoring, \emph{IEEE International Geoscience and Remote Sensing Symposium (IGARSS)}, 2016.
\item S. Dev, Y. H. Lee, S. Winkler, Categorization of Cloud Image Patches Using an Improved Texton-based Approach, \emph{IEEE International Conference on Image Processing (ICIP)}, 2015. 
\item S. Dev, Y. H. Lee, S. Winkler, Multi-level Semantic Labeling of sky/cloud images, \emph{IEEE International Conference on Image Processing (ICIP)}, 2015. 
\item S. Dev, F. M. Savoy, Y. H. Lee, S. Winkler, Design of low-cost, compact and weather-proof whole sky imagers for high-dynamic-range captures, \emph{IEEE International Geoscience and Remote Sensing Symposium (IGARSS)}, 2015.
\item F. M. Savoy, S. Dev, Y. H. Lee, S. Winkler, Cloud base height estimation using high-resolution whole sky imagers, \emph{IEEE International Geoscience and Remote Sensing Symposium (IGARSS)}, 2015.
\item S. Dev, Y. H. Lee, S. Winkler, Systematic Study of Color Spaces and Components for the segmentation of sky/cloud images, \emph{IEEE International Conference on Image Processing (ICIP)}, 2014. 
\item S. Dev, F. M. Savoy, Y. H. Lee, S. Winkler, WAHRSIS: A low-cost, high-resolution whole sky imager with near-infrared capabilities, \emph{Proc. IS\&T/SPIE Infrared Imaging Systems: Design, Analysis, Modeling, and Testing}, 2014.
\end{itemize}

\noindent
\textbf{Technical Reports}
\begin{itemize}[leftmargin=*]
\item S. Dev, F. M. Savoy, Y. H. Lee and S. Winkler, DIY Sky Imager For Weather Observation: A complete guide to build a ground-based sky imager using off-the-shelf components with automatic cloud coverage computation, SPM Student Design Project Series Documents, \emph{IEEE Signal Processing Society}, 2016.\\
(Featured in H. Godrich, Students' Design Project Series: Sharing Experiences [SP Education], IEEE Signal Processing Magazine, Jan 2017.)
\end{itemize}