### 2018-07-16 misc/thesis-hardbound/thesis.tex

\documentclass[reqno,12pt,oneside]{book} % For e-report
%\documentclass[reqno,12pt,twoside]{book} % For printing

\newtheorem{theorem}{Theorem}
\usepackage{longtable}
\usepackage[english]{babel}
\usepackage{fancyhdr}
\usepackage{float} \restylefloat{figure} %defines H option for floats
\usepackage{color}   %May be necessary if you want to color links
\usepackage{setspace}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true, %set true if you want colored links
    linktoc=all,     %set to all if you want both sections and subsections linked
    linkcolor=black,  %choose some color if you want links to stand out
    citecolor=blue,
    urlcolor=blue,
}
\usepackage[nameinlink,noabbrev]{cleveref}
\renewcommand\floatpagefraction{.99}
\renewcommand\dblfloatpagefraction{.95}
\usepackage{graphicx}
\graphicspath{{Figures/}}
\usepackage{multirow}
\usepackage{colortbl}
\usepackage{setspace} %\singlespacing 
\onehalfspacing
\usepackage{fixltx2e}
\usepackage{epstopdf}
\usepackage{rotating}
\usepackage{resizegather}
\usepackage{lscape}
\setlength{\textfloatsep}{1mm}
\setlength{\floatsep}{1mm}
\usepackage{bm} %defines \bm for bold greek letters, i.e. $\bm{\alpha}$
\usepackage{cite}
\usepackage{url}
\usepackage[small]{caption}
\usepackage{subfig} %should come after caption package if given, can't be used with caption2, defines subfloat
\usepackage{array} %More formatting options for array and tabular environments
\usepackage{mathtools}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\usepackage{amsmath}   % From the American Mathematical Society
% A popular package that provides many helpful commands
% for dealing with mathematics. Note that the AMSmath
% package sets \interdisplaylinepenalty to 10000 thus
% preventing page breaks from occurring within multiline
% equations. Use:
\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks
% as IEEEtran.cls normally does. amsmath.sty is already
% installed on most LaTeX systems. The latest version
% and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/
\usepackage{amsfonts}
\usepackage[T1]{fontenc} % For less than and greater than sign
\usepackage{subfloat}
\usepackage{tikz}
\usepackage{afterpage}
\usepackage{lipsum}
\long\def\symbolfootnote[#1]#2{\begingroup\def\thefootnote{\fnsymbol{footnote}}
\footnote[#1]{#2}\endgroup}

\hyphenation{op-tical net-works semi-conduc-tor}

\usepackage{geometry}
\usepackage{enumitem}
\usepackage{stackengine}
\usepackage{wrapfig} % For wrapping figure with text.
\usepackage{algpseudocode} % For algorithm 
\usepackage{algorithm,algpseudocode}

% To hide the references section.
%\usepackage[hide=true]{nobibprint}

\newcolumntype{L}{>{\centering\arraybackslash}m{8mm}} % Used for Table in cloud classification chapter

\definecolor{orange}{rgb}{0.96, 0.52, 0.08}

\newcommand{\forceindent}{\leavevmode{\parindent=1.5em\indent}}

% Addition ends here----------------------

\usepackage{nomencl}

\begin{document}

\frontmatter
\pagestyle{plain}

\begin{titlepage}

\newcommand{\HRule}{\rule{\linewidth}{0.5mm}} % Defines a new command for the horizontal lines, change thickness here

\begin{center}

\textsc{{\Large \bfseries 2D and 3D Image Analysis and Its Application to Sky/Cloud Imaging}}\vspace{0.4cm} % Thesis title

\vspace{4cm}

\textsc{\Large \bfseries Soumyabrata Dev}\\[0.5cm]

\vspace{5cm}

\vfill

\large {School of Electrical \& Electronic Engineering}\\[0.5cm]

\vspace{1cm}

\normalsize{A thesis submitted to the Nanyang Technological University\\ in partial fulfillment of the requirement for the degree of\\ 
Doctor of Philosophy}\\[0.3cm] % University requirement text

{\large \textbf{2017}}\\[4cm] % Date

\end{center}
\end{titlepage}

\cleardoublepage
\input{Misc/ack}
\addcontentsline{toc}{chapter}{Acknowledgment}

\input{Misc/abstract}
\addcontentsline{toc}{chapter}{Abstract}

\clearpage

\pagestyle{fancy}
\fancyhf{}
\afterpage{\cfoot{\thepage}}    %Forces page number in every line

\cleardoublepage
\lhead{\emph{Contents}}
\tableofcontents
\addcontentsline{toc}{chapter}{Contents}

\cleardoublepage
\lhead{\emph{List of Figures}}
\listoffigures
\addcontentsline{toc}{chapter}{\listfigurename}

\cleardoublepage
\lhead{\emph{List of Tables}}
\listoftables
\addcontentsline{toc}{chapter}{List of Tables}

\clearpage

\cleardoublepage

\input{Misc/abbreviations}
\addcontentsline{toc}{chapter}{Abbreviations}
\clearpage

\mainmatter %[This indicates page number will start now]

%INCLUDE THE CHAPTERS HERE
\input{Chapters/Chapter1}
\clearpage

\input{Chapters/Chapter2}
\clearpage

\input{Chapters/Chapter3}
\clearpage

\input{Chapters/Chapter4}
\clearpage

\input{Chapters/Chapter5}
\clearpage

\input{Chapters/Chapter6}
\clearpage

\input{Chapters/Chapter7}
\clearpage

\input{Chapters/Chapter8}
\clearpage

\input{Chapters/Chapter9}
\clearpage

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - APPENDICES
%----------------------------------------------------------------------------------------

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics

\appendix % Cue to tell LaTeX that the following 'chapters' are Appendices
% Suppressing the counter to not show sections and sub-sections.

\addtocontents{toc}{\protect\setcounter{tocdepth}{0}}

% Include the appendices of the thesis as separate files from the Appendices folder
% Uncomment the lines as you write the Appendices

\input{Appendices/AppendixA}
\input{Appendices/AppendixB}
\input{Appendices/AppendixC}
\input{Appendices/AppendixD}

% Setting the counter back on track!
\addtocontents{toc}{\protect\setcounter{tocdepth}{2}}

\backmatter

\cleardoublepage

\input{Misc/notations}
\addcontentsline{toc}{chapter}{Notations}
\clearpage

\input{Misc/pubs}
\addcontentsline{toc}{chapter}{Author's Publications}

%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

\label{Bibliography}
\small
\begin{singlespace}
\lhead{\emph{Bibliography}}
\bibliographystyle{IEEEbib}
\bibliography{./References/longforms,./References/references}	
\end{singlespace}

\end{document}