### 2018-07-16 thesis/1_foreword/ack.tex

\addcontentsline{toc}{chapter}{Acknowledgements}
\begin{center}
\textbf{\large Acknowledgements}
\end{center}

\noindent There are many people I want to thank for helping me in completing this work. Foremost, I would express my sincerest gratitude to my supervisor, Asst. Prof. Chen Shoushun for supporting and guiding me through my work and study with his knowledge and patience. I am very grateful for his insights and instructions in the hardware design for FPGA. \\

\noindent Many thanks to all my labmates of Satellite Research Center II, especially MEHTA Deval Samirbhai and Chin Shi Tong for the first steps of my research and the help whenever I am in troubles and doubt. \\

\noindent I would also like to thank Mrs. Janet Tan and Mrs. Pamela Ng, the technicians of Satellite Research Center II for helping me with all the lab facilities instructions also, setting up the equipment and providing resources whenever I need. \\

\noindent Finally, I sincerely thankful to my parents, my brother and all of my friends for their love and support throughout my study.

\vspace{1cm}

\hfill