### 2018-07-16 thesis/2_chapters/chap5.tex

% Chapter 5
\chapter{Conclusion and Future Works}
\label{chap:conclusion}

\section{Summary}

The purpose of this research topic is to implement a system evaluating a star tracking algorithm proposed by Mr. Minh Duc Pham\cite{MDP,edselc.2-52.0-8487677997120120101,edseee.655799920130101}, my senior and an ex post-graduated NTU student. In my thesis, I have summarized three traditional methods and one state-of-the-art star tracking method proposed in the literature review section. Then in the Hardware and Software Co-processing – Implementation of the Algorithm section, I introduced the hardware system, deeply analyze the algorithm. I also proposed an adaptive algorithm based on the proposed algorithm that is suitable for the software-hardware co-processing implementation in the system. In the final section Experimental Results, I suggested my hardware architecture design for this project, carry out some experiments to evaluate the implementation and their results. \\

\noindent All of my work includes the simulation image generator, image datasets, hardware IP core design, system architecture, software algorithms and source code, results of the experiments, quick documentations are dedicated in this repository: https://github.com/dangkhoadl/Master-Thesis-Star-Tracking-System

\section{Conclusion}

\section{Future Works}