### 2019-04-02 thesis_supports/Response_Reviewer2/Response_Reviewer2.tex

\documentclass[dvips,a4paper,12pt]{report}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage[
    top=0.6in,
    bottom=0.6in,
    right=0.6in,
    left=0.6in]{geometry}

\begin{document}
\pagenumbering{arabic}

% ----------------------------------------- Begin content --------------------------------------------- %
%%% letter
\noindent Respected Examiner,

\noindent Thank you for reviewing my thesis.

\begin{center}
    \textbf{Thesis Title: Implementation of an Autonomous Star Recognition Algorithm using Hardware-Software Co-processing Approach} 
\end{center}

\noindent Firstly, I would like to thank you for your insightful comments. These excellent comments have helped significantly improve my thesis. This letter addresses each of the raised points and notes precisely how I have responded and included the comments into my revised thesis. \\

\vspace{0.5in}

\noindent Sincerely yours,

\noindent Dang Le Dang Khoa

%%% Q&A
\newpage

\color{blue}
\noindent \large{\textbf{Examiner: 2}}

\begin{enumerate}
    \color{blue}
    \item There are many careless mistakes in the thesis. Tables and figures are not explained in the body of the thesis, equations are given without explanations, unfinished sentences, etc. All these are indicated on the relevant pages in the thesis. Please correct and give explanations as requested.

    \color{black}
    Answer

    \color{blue}
    \item Page 29. Please explain why the threshold is chosen to be 70 to separate star clusters from the background. That will happen if the threshold is reduced or increase?

    \color{black}
    Answer

    \color{blue}
    \item Conclusion must be made for the study, so I would like to see a chapter on conclusion which includes the following:
    \begin{enumerate}[label=(\alph*)]
        \color{blue}
        \item One of the objective of the candidate's thesis is to implement and evaluate the performance of a star tracking algorithm, so what is the conclusion of the evaluation of the performance? Can it be used for the attitude determination sensor mounted on a spacecraft?

        \color{black}
        Answer

        \color{blue}
        \item Another objecttive of the thesis is to benchmark and optimize the algorithm in terms of the poer consumption and the area of transistor implementation. What is the conclusion? What is the benchmark now? Did the candidate do optimization? What I got from the thesis are just the results of the power consumed, but it is necessary that the candidate describes how he did the optimization.

        \color{black}
        Answer
    \end{enumerate}

\end{enumerate}

% ----------------------------------------- End content --------------------------------------------- %
\end{document}