### 2019-04-02 thesis/1_foreword/originality.tex

\addcontentsline{toc}{chapter}{Statement of Originality}
\begin{center}
	\textbf{\large Statement of Originality}
\end{center}

\noindent I hereby certify that the work embodied in this thesis is the result of original research, is free of plagiarised materials, and has not been submitted for a higher degree to any other University or Institution. \\

\vspace{2in}

\begin{multicols}{2}
    \begin{flushleft}
        ........................ \\
        \textbf{Date}
    \end{flushleft}

    \columnbreak

    \begin{flushright}
        ........................ \\
        \textbf{Dang Le Dang Khoa} \\
    \end{flushright}
\end{multicols}