### 2019-04-02 thesis/1_foreword/supervisor.tex

\addcontentsline{toc}{chapter}{Supervisor Declaration Statement}
\begin{center}
	\textbf{\large Supervisor Declaration Statement}
\end{center}

\noindent I have reviewed the content and presentation style of this thesis and declare it is free of plagiarism and of sufficient grammatical clarity to be examined. To the best of my knowledge, the research and writing are those of the candidate except as acknowledged in the Author Attribution Statement. I confirm that the investigations were conducted in accord with the ethics policies and integrity standards of Nanyang Technological University and that the research data are presented honestly and without prejudice.

\vspace{2in}

\begin{multicols}{2}
    \begin{flushleft}
        ........................ \\
        \textbf{Date}
    \end{flushleft}

    \columnbreak

    \begin{flushright}
        ........................ \\
        \textbf{Assoc Prof Chen Shoushun} \\
    \end{flushright}
\end{multicols}