### 2019-04-09 thesis/1_foreword/authorship.tex
Add Signed thesis

\addcontentsline{toc}{chapter}{Authorship Attribution Statement}
\begin{center}
    \textbf{\large Authorship Attribution Statement}
\end{center}

\noindent This thesis \textbf{does not} contain any materials from papers published in peer-reviewed journals or from papers accepted at conferences in which I am listed as an author.

\vspace{2in}

\begin{multicols}{2}
    \begin{flushleft}
        ........................ \\
        \textbf{Date}
    \end{flushleft}

    \columnbreak

    \begin{flushright}
        ........................ \\
        \textbf{Dang Le Dang Khoa} \\
    \end{flushright}
\end{multicols}